---
title: "Sorafinib FLT3-ITD Survival"
author: "Jenny Smith"
date: "June 18, 2019"
output: html_document
---

#Set-up

```{r setup}
library(knitr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.align='center', fig.height = 5, fig.width = 7)
knitr::opts_knit$set(root.dir = '/fh/fast/meshinchi_s/workingDir/TARGET/AML_TARGET/Clinical/analysis/2017.10.01_Flt3ITD_Sorafinib/')
options(stringsAsFactors = FALSE)
```


```{r message=FALSE}
library(magrittr)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
getwd()
```

```{r}
source("~/scripts/survival_analysis/Survplot_Functions_2018.10.24.r")
```


#Define Functions to Be Used

```{r}
#https://stackoverflow.com/questions/21724212/set-r-to-include-missing-data-how-can-is-set-the-usena-ifany-option-for-t
table = function (..., useNA = 'ifany') base::table(..., useNA = useNA)
```


#Read in the ITD, and Sorafinib Dosage Information 

itd <- read.csv("~/reference_mapping-files/AAML1031_ITDpos_data_091917.csv", stringsAsFactors = FALSE)
itd <- read.csv("~/reference_mapping-files/AAML1031_ITDpos_data_091917_molecular_added.csv", stringsAsFactors = FALSE)

```{r}
armC.IndI <- read.csv("AAML1031_Arm_C_dose_data_121318_INDUCTION_I.csv")
```

```{r}
#this is the original one used 
armC <- read.csv("AAML1031_ArmC_10.17.2017.csv", stringsAsFactors = FALSE) %>% 
  
  #this is an update version from nearly 2 years later. Need to cross-reference. # it has 7 additional armC patients 
  full_join(., read.csv("AAML1031_Arm_C_dose_data_121318_InductionI.csv"),
            by=c("Pt.ID"="Patient.registration.number")) %>% 
  
  select(Date.of.enrollment_2018=Date.of.enrollment, Treatment.Arm_2018=Treatment.Arm, everything())


dim(armC) #92 patients 
head(armC[,1:5])
```



#Read in the Molecular Data 

```{r}
WT1 <- read.csv("20190215_AAML1031_FLT3_ITD_positive_WT1_results.csv")

head(WT1)
dim(WT1) #177   9
```

```{r}
AR <- read.csv("Copy of AAML1031_FLT3_ITD_less_than_0.1_AR.csv")

head(AR)
```





#Session Information

```{r}
sessionInfo()
```










