---
title: "FLT3-ITD response to sorafinib"
author: "Jenny Smith"
date: ""
output: html_document
---


# Set-up 

```{r setup, cache = FALSE, include = FALSE}
require(knitr)
knitr::opts_knit$set(root.dir = file.path(TARGET,"Clinical/analysis/2017.10.01_Flt3ITD_Sorafinib"))
```

```{r}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),
                      tidy=TRUE,
                      fig.align='center',
                      fig.width = 10, fig.height = 10)
node=Sys.info()[["nodename"]]
if(!grepl("local", node)){
  print(node)
  options(bitmapType = 'cairo')
  grDevices::X11.options(type='cairo')
}

options(stringsAsFactors = FALSE)
table = function (..., useNA = 'ifany') base::table(..., useNA = useNA)
```

```{r message = FALSE, warning=FALSE}
library(stringr)
library(magrittr)

library(ggplot2)
library(gridExtra)

library(dplyr)
library(tidyr)
library(tibble)


library(DeGSEA)
library(REDCapR)
# library(keyring) #cannot install non-zero exit 

getwd()
```


# ClinData

```{r message=FALSE}
#https://cran.r-project.org/web/packages/REDCapR/vignettes/workflow-read.html 
project <- "FLT3-ITD_with_Fusions_Sorafinib_Response"

if(project==""){
  stop("Must include Projecy name!")
}else{
  message(paste0("Project is: ",project))
  current_cde_database <- paste("TARGET_AML_CDEs_For_Project",project, ".RDS", sep="_")
  
  if(file.exists(current_cde_database)){
    print("Reading CDEs from Rdata object.")
    merged <- readRDS(current_cde_database)
      
  }else{
    print("Downloading CDEs from RedCap API.")
    path_credential <- file.path(HOME,".redcap")
    project_id <- 1295
  
    credential  <- REDCapR::retrieve_credential_local(
      path_credential = path_credential,
      project_id = project_id)
    
    #takes about 30 sec to download. 
    merged <- redcap_read(redcap_uri = credential$redcap_uri, 
                          token = credential$token, 
                          raw_or_label_headers = 'label')
    if(merged$success){
      merged <- data.frame(merged$data, check.names=TRUE) #remove the white spaces  (will this even work??)
      saveRDS(merged, current_cde_database)
    }
    
    fname_c <- ifelse(length(ls(pattern = "_cts_file")) == 0,"None" ,basename(get(ls(pattern = "_cts_file"))))
    fname_t <- ifelse(length(basename(ls(pattern = "_TPM_file"))) == 0,"None", basename(get(ls(pattern = "_TPM_file"))))
    
    #Create a simple log file from the day the project starts
    cat(c(paste("Date:", Sys.Date()),
          paste("cts:", fname_c),
          paste("tpm:", fname_t), 
          paste("CDE:", current_cde_database)),
          sep = "\n", 
          file = paste(project, Sys.Date(), ".log", sep="_"))
    
  }
  
  #keep a list of the ineligable patiens to remove if necessary
  inelig <- merged %>% 
    filter(Eligibility.Comments == "remove") %>% 
    pull(USI)
  
  #Filter those with USIs and are eligible for the study
  merged <- merged %>% 
    filter(Eligibility.Comments != "remove")
  
  
  dim(merged) 
  head(merged)

}
```


```{r}
sample_info <- read.csv(file.path(TARGET, "SequencingDataMatrix/TARGET_AML_Ribodepleted_Manifest_08.12.21.csv")) 

dim(sample_info)
```

```{r}
regs <- read.csv(file.path(CDE,"Merged/TARGET_AML_0531_1031_merged_CDEs_05.21.21.csv")) %>% 
  select(Reg., USI)
```


```{r}
sorafinib_data <- read.csv("AAML1031_ArmC_10.17.2017.csv") %>% 
  mutate(Pt.ID=as.character(Pt.ID)) %>% 
  left_join(., regs, by=c("Pt.ID"="Reg.")) %>% 
  left_join(., select(merged,Primary.Fusion,FLT3.ITD.positive., USI),
            by="USI") %>% 
  select(USI, Pt.ID, 
         Primary.Fusion, FLT3.ITD.positive.,
         everything())

head(sorafinib_data)
```


# select Samples

Do you already have an outcome curve created for AAML1031 patients treated with sorafenib (ITD pos) comparing NUP98-NSD1 and DEK-NUP214?  Soheil would like to include in his talk that essentially sorafenib doesn’t benefit these 2 ITD cohorts.  If not, could you create one?  Just 2 populations needed: ITD+ NSD1 (treated with sorafenib) compared to ITD+ DEK-NUP214 (treated with sorafenib) for OS.

can you create one with all 6;9 patients regardless on arm c or not?

```{r}
table(merged$FLT3.ITD.positive.)
```

```{r}
input_samples <- merged %>% 
  filter(grepl("NSD1|DEK-NUP214", Primary.Fusion), 
         grepl("Yes", FLT3.ITD.positive.)) %>% 
  filter(Protocol=="AAML1031") %>% 
    filter( grepl("DEK-NUP214", Primary.Fusion) | 
              c(grepl("Arm C", Treatment.Arm) & Primary.Fusion == "NUP98-NSD1")) %>% 
  rename_at(vars(c(Recoded.OS.ID, Recoded.Event.ID)),
            ~c("OS.ID","Event.ID")) %>% 
  filter(!is.na(EFS.time..days.))


table(input_samples$EFS.event.type.ID)
```

```{r}
input_samples %>% 
  group_by(Primary.Fusion, FLT3.ITD.positive., Treatment.Arm) %>% 
  dplyr::count() #only 3 DEK-NUP214 on the sorafinib arm?? 

input_samples %>% 
  group_by(Primary.Fusion, EFS.event.type.ID) %>% 
  dplyr::count() #only 3 DEK-NUP214 on the sorafinib arm??
```

```{r}
orig <- sorafinib_data %>% 
        filter(grepl("NSD1|DEK-NUP214", Primary.Fusion), 
               grepl("Yes", FLT3.ITD.positive.))


table(orig$Primary.Fusion) #appears to be that way - even in the original dataset so unlikey any were lost to ineligibility, or other updates. 
``` 

```{r}
aaml0531_samples <- merged %>% 
  filter(grepl("NSD1|DEK-NUP214", Primary.Fusion), 
         grepl("Yes", FLT3.ITD.positive.)) %>% 
  filter(Protocol=="AAML0531") %>% 
  # filter( grepl("DEK-NUP214", Primary.Fusion) | 
  #             c(grepl("Arm C", Treatment.Arm) & Primary.Fusion == "NUP98-NSD1")) %>% 
  rename_at(vars(c(Recoded.OS.ID, Recoded.Event.ID)),
            ~c("OS.ID","Event.ID")) %>% 
  filter(!is.na(EFS.time..days.))



aaml0531_samples %>% 
  group_by(Primary.Fusion,FLT3.ITD.positive., Treatment.Arm) %>% 
  dplyr::count() #only 3 DEK-NUP214 on the sorafinib arm?? 

aaml0531_samples %>% 
  group_by(Primary.Fusion, EFS.event.type.ID) %>% 
  dplyr::count() #only 3 DEK-NUP214 on the sorafinib arm??
```

# KM plots 

```{r}
source("~/home_drive/scripts/RNAseq_Analysis/DifferentialExpn_PathwayAnalysis/R/Survplot_Functions_2018.10.24.r")
```


```{r}
NSD1_and_DEKs <- KM.plots(df=input_samples,
                          group_vars = NULL,
                          type = "OS", 
                          covariate = "Primary.Fusion",cohort = "1031",
                          cc = c("NUP98-NSD1"="red", "DEK-NUP214"="navy"))

NSD1_and_DEKs
```

```{r}
# pdf("TARGET_AML_Arm_C_NUP98.NSD1_and_Any_DEK.NUP214_OS_KMplot.pdf", height = 6, width = 6)
grid.arrange(grobs=c(NSD1_and_DEKs$OS))
# dev.off()

# pdf("TARGET_AML_Arm_C_NUP98.NSD1_and_Any_DEK.NUP214_EFS_KMplot.pdf", height=6, width = 6)
grid.arrange(grobs=c(NSD1_and_DEKs$EFS))
# dev.off()
```

```{r}
NSD1_and_DEKs_0531 <- KM.plots(df=aaml0531_samples,
                          group_vars = NULL,
                          type = "OS", 
                          covariate = "Primary.Fusion",
                          cohort = "1031",
                          cc = c("NUP98-NSD1"="red", "DEK-NUP214"="navy"))

NSD1_and_DEKs_0531
```

```{r}
# pdf("TARGET_AAML0531_FLT3-ITD+_NUP98.NSD1_and_Any_FLT3-ITD+_DEK.NUP214_OS_KMplot.pdf", height = 6, width = 6)
# grid.arrange(grobs=NSD1_and_DEKs_0531$OS)
# dev.off()
# 
# pdf("TARGET_AAML0531_FLT3-ITD+_NUP98.NSD1_and_Any_FLT3-ITD+_DEK.NUP214_EFS_KMplot.pdf", height = 6, width = 6)
# grid.arrange(grobs=NSD1_and_DEKs_0531$EFS)
# dev.off()
```


#Session Information

```{r}
sessionInfo()
```

